```{r import_libraries, warning=FALSE, message=FALSE, }
library(foreign) # read.spss
library(psych) # describe 
library(Hmisc) # rcorr 
library(QuantPsyc) # lm.beta
library(car) # vif, durbinWatsonTest
library(MASS) # studres
library(lmSupport) #lm.sumSquares
library(perturb) # colldiag
```

In order to emulate SPSS output, it is necessary to install several add-on packages. The above `library` commands load the packages into your R workspace. I've highlighted in the comment the names of the functions that are used in this script. 

You may not have the above packages installed.
If not, run commands like:
* `install.packages('foreign')`
* `install.packages('psych')`
* etc.

for each of the above packages not installed or use the "packages" tab in RStudio to install.

Note also that much of this analysis could be performed using [Rcommander](http://www.rcommander.com/) using a more SPSS-style GUI environment.

# Import and prepare data
```{r import_data}
cars_raw <- read.spss('cars.sav', to.data.frame=TRUE)
# get rid of missing data listwise
cars <- na.omit(cars_raw[,c('accel', 'mpg', 'engine', 'horse', 'weight')])
```

Ensure that `cars.sav` is the working directory.

# Quick look at data
```{r quick_look}
# note the need to deal with missing data
psych::describe(cars_raw)

dim(cars)
head(cars)
str(cars)
```

# Fit model
```{r}
fit <- lm(accel ~ mpg + engine + horse + weight, data=cars)
```

## Descriptive Statistics
```{r}
# Descriptive statistics
psych::describe(cars)

# correlations
cor(cars)
rcorr(as.matrix(cars)) # include sig test for all correlations
# scatterplot matrix if you want
pairs.panels(cars)
```

## Summary of model
```{r}
# r-square, adjusted r-square, std. error of estimate,
# overall ANOVA, df, p, 
# unstandardised coefficients, sig tests
summary(fit)
### additional info in terms of sums of squares
anova(fit)

# 95% confidence intervals (defaults to 95%)
confint(fit)
# but can specify different confidence intervals
confint(fit, level=.99)

# standardised coefficients
lm.beta(fit)

# or you could do it manually
zcars <- data.frame(scale(cars)) # make all variables z-scores
zfit <- lm(accel ~ mpg + engine + horse + weight, data=zcars)
coef(zfit)[-1]

# correlations: zero-order, semi-partial, partial
# obscure function seems to do it
sqrt(lm.sumSquares(fit)[,c(2,3)])

# or use own function 
cor_lm <- function(fit) {
    dv <- names(fit$model)[1]
    dv_data <- fit$model[,dv]
    ivs <- names(fit$model)[-1]
    iv_data <- fit$model[, ivs]
    x <- fit$model
    x_omit <- lapply(ivs, function(X) x[,c(dv, setdiff(ivs, X))])
    names(x_omit) <- ivs
    lapply(x_omit, head)
    fits_omit <- lapply(x_omit, function(X) 
        lm(as.formula(paste(dv, "~ .")), data=X))
    resid_omit <- sapply(fits_omit, resid)
    iv_omit <- lapply(ivs, function(X) 
        lm(as.formula(paste(X, "~ .")), data=iv_data))
    resid_iv_omit <- sapply(iv_omit, resid)
    
    results <- sapply(seq(ivs), function(i) 
        c(zeroorder=cor(iv_data[,i], dv_data), 
          partial= cor(resid_iv_omit[,i], resid_omit[,i]),
          semipartial=cor(resid_iv_omit[,i], dv_data) 
          ))
    results <- data.frame(results)
    
    names(results) <- ivs
    results <- data.frame(t(results))
    results
}
        
round(cor_lm(fit), 3)
```

## Assumption testing
```{r}
# Durbin Watson test
durbinWatsonTest(fit)

# vif
vif(fit)

# tolerance
1/vif(fit)

# collinearity diagnostics
colldiag(fit)

# residual statistics
rfit <- data.frame(predicted=predict(fit),  
           residuals=resid(fit), 
           studentised_residuals=studres(fit))
psych::describe(rfit)

# distribution of standarised residuals
zresid <- scale(resid(fit))
hist(zresid)
# or add normal curve
# http://www.statmethods.net/graphs/density.html 
hist_with_normal_curve <- function(x, breaks=24) {
    h <- hist(zresid, breaks=breaks, col='lightblue')
    xfit<-seq(min(x),max(x),length=40) 
    yfit<-dnorm(xfit,mean=mean(x),sd=sd(x)) 
    yfit <- yfit*diff(h$mids[1:2])*length(x) 
    lines(xfit, yfit, lwd=2)
}
hist_with_normal_curve(zresid)

# normality of residuals
qqnorm(zresid)
abline(a=0, b=1)

# plot predicted by residual 
plot(predict(fit), resid(fit))

# plot dependent by residual
plot(cars$accel, resid(fit)) 
```

